Genetic alterations and expression patterns of genes coding for FGF, FGFR, and FGFBP proteins in urothelial cancers

### 2024-12-10 report/markdown/report.Rmd
Figure with mutation hot spots
And an update of the scheme of FGFR signaling

* Mutations of _FGFR1_, _FGFR2_, _FGFR3_, and _FGFR4_ were predominantly classified as missense mutations caused by single nucleotide polymorphisms (__Figure \@ref(fig:fig-mutation-type)__ and __\@ref(fig:fig-variant-type)__, __Table \@ref(tab:tab-fgfr-class)__).  For _FGFR1_, the mutation hot-spots were located in the Ig-like domain 1 and kinase domain. The kinase domain was the prime target of mutations of _FGFR2_. For _FGFR3_, they affected protein residues were located between Ig-like domains 1 and 2 responsible for ligand biding, or at the membrane-spanning hinge between the ligand-binding region and the kinase domain responsible for activation of the downstream signaling. In case of _FGFR4_, the Ig-like domain 1, the hinge region between the Ig-like domain 3 and kinase domain, as well as the kinase domain were prime targets of mutations of _FGFR4_ (__Figures \@ref(fig:fig-protein-domain-split)
__ - __\@ref(fig:fig-fgfr4-position)__, __Table \@ref(tab:tab-fgfr-class)__). It can be hence speculated, that somatic mutations of the _FGFR_ genes may affect both the kinase activity and the conformation of the protein resulting in a higher affinity towards ligands or ligand-independent signaling.

@@

>>>>

* Mutations of _FGFR1_, _FGFR2_, _FGFR3_, and _FGFR4_ were predominantly classified as missense mutations caused by single nucleotide polymorphisms (__Figure \@ref(fig:fig-mutation-type)__ and __\@ref(fig:fig-variant-type)__, __Table \@ref(tab:tab-fgfr-class)__).  For _FGFR1_, the mutation hot-spots were located in the Ig-like domain 1 and kinase domain. The kinase domain was the prime target of mutations of _FGFR2_. For _FGFR3_, they affected protein residues were located between Ig-like domains 1 and 2 responsible for ligand biding, or at the membrane-spanning hinge between the ligand-binding region and the kinase domain responsible for activation of the downstream signaling. In case of _FGFR4_, the Ig-like domain 1, the hinge region between the Ig-like domain 3 and kinase domain, as well as the kinase domain were prime targets of mutations of _FGFR4_ (__Figures \@ref(fig:fig-protein-domain-split)
__ - __\@ref(fig:fig-fgfr4-position)__, __Table \@ref(tab:tab-fgfr-class)__). It can be hence speculated, that somatic mutations of the _FGFR_ genes may affect both the kinase activity and the conformation of the protein resulting in a higher affinity towards ligands or ligand-independent signaling. In a more detailed analysis of mutation hot spots of the _FGFR3_ gene, mutation resulting in changes of protein residues S249C (`r expl_hot$frequency$protein_domain$FGFR3 %>% filter(protein_change == 'S249C') %>% .$percent %>% signif(2) %>% range %>% paste(collapse = ' to ')`% of cancer samples), Y373C (`r expl_hot$frequency$protein_domain$FGFR3 %>% filter(protein_change == 'Y373C') %>% .$percent %>% signif(2) %>% range %>% paste(collapse = ' to ')`%), and R248C (`r expl_hot$frequency$protein_domain$FGFR3 %>% filter(protein_change == 'R248C') %>% .$percent %>% signif(2) %>% range %>% paste(collapse = ' to ')`%) were present in $\geq$ 1% of cancer samples in at least two out of the `r globals$cohort_labs[c("genie", "msk", "tcga", "bcan")] %>% collapse_and` cohorts (__Figure \@ref(fig:fig-fgfr3-hotspots)__).

@@

```{r fig-fgfr3-hotspots, fig.width = figur::convert(rep_figs$fgfr3_hotspots, to = 'in')$w, fig.height = figur::convert(rep_figs$fgfr3_hotspots, to = 'in')$h, fig.cap = 'Mutational hot spots of the FGFR3 gene.'}

rep_figs$fgfr3_hotspots$plot

__Figure \@ref(fig:fig-fgfr3-hotspots). Mutational hot spots of the FGFR3 gene.__

_(A) Density of mutations along the FGFR3 protein sequence was calculated with the sliding window method, scaled to the percentage within all cancer samples in the cohort, and displayed in dot plots. Each point represents a single protein residue, point color codes for assignment of the residue to the protein domain. Numbers of analyzed cancer samples are indicated in the plot facets. The protein domain scheme is displayed above the plot._

_(B) Frequency of mutations affecting particular residues of the FGFR3 protein was expressed as percentage of all cancer samples in the cohorts and displayed in dot plots. Each point represents a single protein residue, point color codes for assignment of the residue to the protein domain. Candidate mutation hot spots with mutation rates $\geq$ 1% are highlighted with enlarged symbols, and labeled with percentages and protein sequence changes. Numbers of analyzed cancer samples are indicated in the plot facets. The protein domain scheme is displayed above the plot._
